---
title: "Lab 6"
author: "Yong Jin Nam"
date: "10/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/UMich_Bio201_F19/")
```

# Load packages
```{r Load packages, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(readxl)
library(broom)
library(cowplot)
library(agricolae)
set.seed(7)
```

# Import data

### Import + format SCFA data

Import the SCFA data frame from last week. Update the columns names to match the course's style, set the order of the semesters by converting from character to factor. 

```{r include=FALSE}
scfa_long <- read_delim("Lab5/curated_data/SCFA_wkly_long.txt", 
                        delim = "\t", col_names = TRUE, trim_ws = TRUE, na = c("NA")) %>%
  rename_all(tolower) %>%
  select(-ends_with("median")) %>% 
  mutate(totalscfa_mean = acetate_mean + butyrate_mean + propionate_mean,
         semester = factor(semester,
                           levels=c("Winter2015", "Fall2015", 
                                    "Winter2016", "Fall2016", 
                                    "Winter2017", "Fall2017", 
                                    "Winter2018", "Fall2018", 
                                    "Winter2019"), ordered = TRUE)) %>%
  filter(quantity_compliant == "yes")
```

### Import tax data 

Import the taxonomic table `person_week_mean_by_cat_w_phylum.csv`. Notice the extension is csv, the delimitor in this file is a comma instead of a tab (which is different from what we have worked with previously). 
```{r}
# taxanomic data
taxa_category <- read_delim("Lab6/raw_data/person_week_mean_by_cat_w_phylum.csv",
                            delim = ",", escape_double = FALSE, 
                            trim_ws = TRUE, na=c("NA"),
                            col_types = list())
View(taxa_category)
```

Look at the data frame, notice the week and semester formats are different then the rest of our data files. Today we are going to do some correlations between taxon abundance and short chain fatty acid (SCFA) concentrations. If we want to join these two data sets we will need make sure the format of these variables are the same. 

### Format tax data
```{r}
# create duplicate of data frame
taxa_cat_curated <- taxa_category %>%
  rename_all(tolower) %>%
  rename(participant_id = person,
         study_week = week,
         relative_abundance = fract) %>%
  select(participant_id, study_week, semester, tax_id, relative_abundance, phylum, tax_name, category)

# update week format 
taxa_cat_curated[["study_week"]] <- recode(.x=taxa_cat_curated[["study_week"]], 
                                  "week_1"="week1")
taxa_cat_curated[["study_week"]] <- recode(.x=taxa_cat_curated[["study_week"]], 
                                  "week_3"="week3")
# update semester format
taxa_cat_curated[["semester"]] <- recode(.x=taxa_cat_curated[["semester"]], 
                                  "2015w"="Winter2015")
taxa_cat_curated[["semester"]] <- recode(.x=taxa_cat_curated[["semester"]], 
                                  "2015f"="Fall2015")
taxa_cat_curated[["semester"]] <- recode(.x=taxa_cat_curated[["semester"]], 
                                  "2016w"="Winter2016")
taxa_cat_curated[["semester"]] <- recode(.x=taxa_cat_curated[["semester"]], 
                                  "2016f"="Fall2016")
taxa_cat_curated[["semester"]] <- recode(.x=taxa_cat_curated[["semester"]], 
                                  "2017w"="Winter2017")
taxa_cat_curated[["semester"]] <- recode(.x=taxa_cat_curated[["semester"]], 
                                  "2017f"="Fall2017")
taxa_cat_curated[["semester"]] <- recode(.x=taxa_cat_curated[["semester"]], 
                                  "2018w"="Winter2018")
taxa_cat_curated[["semester"]] <- recode(.x=taxa_cat_curated[["semester"]], 
                                  "2018f"="Fall2018")
taxa_cat_curated[["semester"]] <- recode(.x=taxa_cat_curated[["semester"]], 
                                  "2019w"="Winter2019")

# Save a copy of the curated data frame
write_delim(taxa_cat_curated, path="Lab6/curated_data/taxa_cat_curated.csv",
            delim = ",", quote = FALSE, col_names = TRUE)

View(taxa_cat_curated)
```

The cleaned-up `taxa_cat_curated` contains a subset of organisms found in the gut, specifically organisms which consume or produce metabolites of interest. The `tax_id` column is a number assigned to a sequence during the sequence processing pipeline. You won't need to use these right now. The `relative_abundance` column contains the frequencey of the organisms in each participant. The frequency is the weekly mean of (# sequences for each tax_id) / (total sequences per sample) for each participant. The `tax_name` is the most likely identity of each sequence. Information regarding microbial metabolism is in the `category` column:

* RS_degraders = Resistant starch degrading specialists (Bifidobacterium + Ruminococcus bromii)
* butyrogen_focused = A focused list of the most common / canonical butyrogens
* butyrogen_flint = A broader list of butyrogenic organisms (information comes from Harry Flint's research)
* hydrogen_consumers = Hydrogen consumers
* hydrogen_gen = Hydrogen generators
* propionate_gen = Propionate generators (essentially Bacteroides)

### Join SCFA + tax data

In the latter half of this lab we are going to test correlations. To test correlations between taxa and their metabolites, we need to join the data frames containing these data. Join the `tax_cat_curated` and `scfa_long` data frames. Remember to check the number of rows and participants in each data frame to help you select the best type of join. Discuss your decision with your neighbor, did you choose the same type of join? 

```{r}
#check number of rows
nrow(taxa_cat_curated) 
nrow(scfa_long) 

#check number of participants 
n_distinct(taxa_cat_curated$participant_id) 
n_distinct(scfa_long$participant_id) 

#identify shared columns = ?
colnames(taxa_cat_curated) 
colnames(scfa_long)
```

```{r}
tax_scfa <- inner_join(scfa_long, taxa_cat_curated,
                 by = c())
```

# Intro

Last week we discussed the factors influencing data structure (sample size, varaince), how to check assumptions for statistical tests, how to determine if the differences between two means (t-tests, Mann-Whitney-U) or more (ANOVA) are likely real, and what to do if your datset does not meet the assumptions when comparing two means (non-parametric). 

In the first portion of the lab we will review the assumption checking process for comparison of more than 2 means, use the non-pararmetric counterpart to ANOVA (Kruskall-Wallace), and learn what to do if your ANOVA (or Kruskall-Wallace) test has a significant result. This will complete our section on categorical/discrete vs. continuous variables. The second half of the lab will discuss how to plot and analyze continuous vs. countinuous variables.

### Checking assumptions

Remember the assumptions for an ANOVA are similar to a t-test (large-ish sample size, normal distribution, equal variances). 

```{r eval=FALSE, include=FALSE}
# sample size of each group
group_by() %>% summarize(sample_size = n())

# check normality of each group
shaprio.test()
histogram()

# check variances
bartlett.test()
```

Last week each ANOVA you conducted should have resulted in a non-significant result (approx p >= 0.15). Run the code for the plot and ANOVA below. 

```{r}
# know assumptions are already met
scfa_long %>%
  filter(study_week == "week1") %>%
  ggplot(aes(x = semester,
             y = butyrate_mean, 
             color = semester)) +
  geom_violin() + 
  geom_jitter() + 
  xlab(NULL) +
  ylab("Week 1 butyrate (mmol/kg)") + 
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# ANOVA 
aov_wk1 <- aov(butyrate_mean ~ semester, scfa_long)
summary(aov_wk1)
```

Here we have a small p-value, this means at least one of our groups is significantly different from the others. To identify which group is different we have to run a pos-hoc test (an additional test). We will use Tukey's Honestly Significant Difference (HSD).

```{r}
# conduct Tukey's HSD test 
tuk_wk1 <- tidy(TukeyHSD(aov_wk1)) %>%
  mutate(adj.p.value = round(adj.p.value, digits = 2)) %>%
  filter(adj.p.value <= 0.10) #filter Tukey output for significant results 

# save data frame of results to file 
write_delim(tuk_wk1, path = "~/Documents/UMich_Bio201_F19/Lab6/curated_data/tuk_wk1.txt",
            delim = "\t", col_names = TRUE, quote = FALSE)
```

### Kruskall-Wallace

If you have a data set that is clearly non-normal (such as the breath methane data), and want to ask a similar question as we asked with ANOVAs last week (e.g., Does the breath methane differ during week 3 for individuals consuming different supplements?), we need to use a non-parametric test. The non-parametric companion to an ANOVA is the Kruskall-Wallace test. 

Last week you should have checked assumptions on breath methane and determined that it was non-normal, so an ANOVA would not be an appropriate test to use to determine if the week 3 breath methane concentrations differed between supplements consumed.

```{r}
# import methane data 
breath_wkly <- read_delim("~/Documents/UMich_Bio201_F19/Lab5/curated_data/breath_wkly.txt", 
                        delim = "\t", col_names = TRUE, trim_ws = TRUE, na = c("NA")) %>%
# format methane data
  rename_all(tolower) %>%
  filter(study_week == "week3",
         quantity_compliant == "yes") %>% #remove non-compliant individuals
  mutate(semester = factor(semester,
                           levels=c("Winter2015", "Fall2015", 
                                    "Winter2016", "Fall2016", 
                                    "Winter2017", "Fall2017", 
                                    "Winter2018", "Fall2018", 
                                    "Winter2019"), ordered = TRUE))
# plot
breath_wkly %>%
  ggplot(aes(x = supplement_consumed, 
             y = ch4_mean, 
             color = supplement_consumed)) + 
  geom_violin() + 
  geom_jitter() + 
  xlab(NULL) + 
  ylab("Breath methane (ppm)") + 
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

# run the test
kruskal.test(supplement_consumed ~ ch4_mean, breath_wkly)
```

The interpretation for Kruskall-Wallace is similar to an ANOVA; small p-value indicates at least one group is different from the others. 

### Homework 6.1

* Write hypotheses to answer the question; do hydrogen breath levels differ between participants consuming fiber supplements? Create a plot of hydrogen levels:
```{r}
# hypothesis: Depending on which supplement was consumed, the hydrogen breath levels will differ because abundance of certain supplements increases abundance of hydrogen producing organisms.

# plot 
breath_wkly %>%
  ggplot(aes(x = supplement_consumed,
             y = h2_mean,
             color = supplement_consumed)) +
  geom_violin() + 
  geom_jitter() + 
  xlab(NULL) + 
  ylab("Breath hydrogen (ppm)") + 
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
             
  
```

* Which assumptions need checked for the statistical test(s)? Check the assumptions in the space below:
```{r}
# check sample size
breath_wkly %>%
  group_by(supplement_consumed) %>% summarize(sample_size = n())

# check normality 
b1 <- breath_wkly %>%
  filter(supplement_consumed == "Accessible")
shapiro.test(b1$h2_mean)
ggplot(b1, aes(x=h2_mean)) +
  geom_histogram()

b2 <- breath_wkly %>%
  filter(supplement_consumed == "Arabino")
shapiro.test(b2$h2_mean)
ggplot(b2, aes(x=h2_mean)) +
  geom_histogram()

b3 <- breath_wkly %>%
  filter(supplement_consumed == "Banana")
shapiro.test(b3$h2_mean)
ggplot(b3, aes(x=h2_mean)) +
  geom_histogram()

b4 <- breath_wkly %>%
  filter(supplement_consumed == "BRMPS")
shapiro.test(b4$h2_mean)
ggplot(b4, aes(x=h2_mean)) +
  geom_histogram()

b5 <- breath_wkly %>%
  filter(supplement_consumed == "BRMPS+Accessible")
shapiro.test(b5$h2_mean)
ggplot(b5, aes(x=h2_mean)) +
  geom_histogram()

b6 <- breath_wkly %>%
  filter(supplement_consumed == "HiMaize")
shapiro.test(b6$h2_mean)
ggplot(b6, aes(x=h2_mean)) +
  geom_histogram()
  
b7 <- breath_wkly %>%
  filter(supplement_consumed == "HiMaize+BRMPS")
shapiro.test(b7$h2_mean)
ggplot(b7, aes(x=h2_mean)) +
  geom_histogram()

b8 <- breath_wkly %>%
  filter(supplement_consumed == "Inulin")
shapiro.test(b8$h2_mean)
ggplot(b8, aes(x=h2_mean)) +
  geom_histogram()

b9 <- breath_wkly %>%
  filter(supplement_consumed == "LOODAT")
shapiro.test(b9$h2_mean)
ggplot(b9, aes(x=h2_mean)) +
  geom_histogram()

b10 <- breath_wkly %>%
  filter(supplement_consumed == "Psyllium")
shapiro.test(b10$h2_mean)
ggplot(b10, aes(x=h2_mean)) +
  geom_histogram()

b11 <- breath_wkly %>%
  filter(supplement_consumed == "Psyllium+BRMPS")
shapiro.test(b11$h2_mean)
ggplot(b11, aes(x=h2_mean)) +
  geom_histogram()

b12 <- breath_wkly %>%
  filter(supplement_consumed == "transition_HiMaize")
shapiro.test(b12$h2_mean)
ggplot(b12, aes(x=h2_mean)) +
  geom_histogram()

# check variance
bartlett.test(h2_mean ~ supplement_consumed, data = breath_wkly)





```

* Which statistical test(s) would you use to answer the question? Conduct the appropriate statistical test(s):
```{r}
kruskal.test(supplement_consumed ~ h2_mean, breath_wkly)
```

* What is the answer to the question? Write your conclusions:
```{r}
# conclusion: The Kruskal-Wallis test produced a p-value of 0.3001, which is relatively high, meaning that there is not enough evidence to conclude that at least one group is different from the others. The hydrogen breath levels do not seem to differ due to the supplement consumed.

```

# Correlations

Correlations are used to determine if there is a relationship between two continuous variables. For the questions we are interested in, determining if there is a monotonic (or linear) relationship will provide insight. A monotonic relationship would occur when both continuous variables increase or decrease simultaneously, or when one variable increases the other decreases. 

To test correlations in this course, we are going to use the best practice approach of linear models (lm). However we do want to point out other correlation tests in case you encounter them elsewhere (Pearson which is parametric and Spearman Rank with is non-parametric). These tests have assumptions which are easy to violate, and the statistics provided (p and R values) do not have a straightfoward interpretation.

### Plot 2 continuous variables

To plot two continuous variables we will use a combination of geom_point() and geom_smooth(). 

```{r}
tax_scfa %>%
  filter(category == "butyrogen_focused",
         study_week == "week3",
         supplement_consumed == "HiMaize") %>%
  ggplot(aes(x = butyrate_mean,
             y = relative_abundance)) + 
  geom_point() + #puts data points to match x and y coordinates
  geom_smooth(method = "lm", #used to create a linear best fit line
              se = FALSE) + #hides confidence interval around line 
  xlab("Week 3 mean butyrate (mmol/kg)") + 
  ylab("Relative abundance") 
```

You plot two continuous variables and they appear to be correlated, you can determine if they are correlated by building a linear model. 

### Linear models

The linear model function is lm(). The linear models we will use have similar arguments as running an ANOVA. You must have the two variables (x and y) and a data frame in which each variable is a column. To access the results of the model, assign it to a new object. In the example below we will determine if there is a correlation between butyrate concentration and the abundance of some butyrate producing organisms. 

```{r}
m1 <- tax_scfa %>%
  filter(category == "butyrogen_focused", #subset dataframe to butyrate producers
         study_week == "week3", #subset for week3
         supplement_consumed == "HiMaize") %>% #subset for HiMaize consumers 
  lm(butyrate_mean ~ relative_abundance, data = .) #test relationship
summary(m1) #view results 
```

Look at the p-value in the same line as the F-statistic, this tells us if there is a significant result. This p-value is high (p-value = 0.2959), so we do not have significance. However if it was significant, we would then look at the adjusted R-squared value which indicates the quality of linear fit (higher R-squared is better >=0.8).

### Homework 6.2

Use the `tax_scfa` data frame, calculate the ratio of each SCFA in a new column (scfa/total). Check that your ratios sum to 1. 

```{r}
scfa_ratio <- tax_scfa %>%
  mutate(acetate_ratio = acetate_mean / totalscfa_mean) %>% 
  mutate(butyrate_ratio = butyrate_mean / totalscfa_mean) %>% 
  mutate(propionate_ratio = propionate_mean / totalscfa_mean) %>%
  mutate(ratio_check = acetate_ratio + butyrate_ratio + propionate_ratio)
```

Calculate the ratios of Firmicutes to Bacteroidetes (Firmicutes abund / Bacteroidetes abundance).

```{r}
# filter for Firmicutes, calculate sum per person per week (firmicutes_abund)
Fr <- scfa_ratio %>%
  filter(phylum == "Firmicutes") %>%
  group_by(participant_id, study_week) %>%
  summarize(firmicutes_abund = sum(relative_abundance))

# filter for Bacteroidetes, calculate sum per person per week (bacteroidetes_abund)
Br <- scfa_ratio %>%
  filter(phylum == "Bacteroidetes <phylum>") %>%
  group_by(participant_id, study_week) %>%
  summarize(bacteroidetes_abund = sum(relative_abundance))

# join subsetted data frames
tax_ratio <- full_join(Fr, Br, by = c()) %>%
  mutate(f_b_ratio = (firmicutes_abund / bacteroidetes_abund)) %>% #calculate ratios
  inner_join(scfa_ratio, tax_ratio, by = c()) #join with scfa_ratio data frame
```

### Homework 6.3
Determine if there is a relationship between the butyrate concentration and Firmicutes abundance. Conduct the statistical test and generate a plot. Save plot to folder within lab 6 called 'figures'.
```{r}
# plot
but_fr_plot <- tax_ratio %>%
  ggplot(aes(x = butyrate_mean,
             y = firmicutes_abund)) +
  geom_point() + 
  geom_smooth(method = "lm", 
              se = FALSE)

save_plot(filename = "Lab6/figures/but_fr_plot.pdf",
          plot = but_fr_plot,
          nrow = 1, ncol = 2,
          base_aspect_ratio = 1.1)

but_fr_plot

# correlation test 
m63a <- tax_ratio %>%
  lm(butyrate_mean ~ firmicutes_abund, data = .)

summary(m63a) # The p-value is low, so we look at the adjusted R-squared value, which is significantly low as well (0.004). There seems to be no strong linear correlation between butyrate concentration and Firmicutes abundance. 
```

Do you draw the same conclusions if you use the butyrate ratio and Firmicutes to Bacteroidetes ratio (f_b_ratio)?
```{r}
# plot
but_f_b_plot <- tax_ratio %>%
  ggplot(aes(x = butyrate_ratio,
             y = f_b_ratio)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE)

save_plot(filename = "Lab6/figures/but_f_b_plot.pdf",
          plot = but_f_b_plot,
          nrow = 1, ncol = 2, 
          base_aspect_ratio = 1.1)

but_f_b_plot

# correlation test 
m63b <- tax_ratio %>%
  lm(butyrate_ratio ~ f_b_ratio, data= .)

summary(m63b) # yes, there still seems to be no strong linear correlation between butyrate ratio and Firmicutes to Bacteroidetes ratio, as shown in the very low adjusted R-squared value (0.03)
```

### Homework 6.4
Determine if there is a relationship between the propionate concentration and Bacteroidetes abundance. Conduct the statistical test and generate a plot. Save the plot.
```{r}
pro_br_plot <- tax_ratio %>%
  ggplot(aes(x = propionate_mean,
             y = bacteroidetes_abund)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE)

save_plot(filename = "Lab6/figures/pro_br_plot.pdf",
          plot = pro_br_plot,
          nrow = 1, ncol = 2,
          base_aspect_ratio = 1.1)

pro_br_plot

m63c <- tax_ratio %>%
  lm(propionate_mean ~ bacteroidetes_abund, data = .) 

summary(m63c) # There seems to be no strong correlation between propionate concentration and Bacteroidetes abundance. Although the p-value produced by the test was low, the adjusted R-squared value was also very low (0.04), indicating a very weak linear relationship between the two variables.
```

Do you draw the same conclusions if you use the propionate ratio and Firmicutes to Bacteroidetes ratio?
```{r}
pro_br_plot2 <- tax_ratio %>%
  ggplot(aes(x = propionate_ratio,
             y = f_b_ratio)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE)

save_plot(filename = "Lab6/figures/pro_br_plot2.pdf",
          plot = pro_br_plot2,
          nrow = 1, ncol = 2,
          base_aspect_ratio = 1.1)

pro_br_plot2

m63d <- tax_ratio %>%
  lm(propionate_ratio ~ f_b_ratio, data = .) 

summary(m63d) # p-value is low but the adjusted R-squared value (0.02) still does not indicate that the linear fit is good. There does not seem to be a relationshipbetween propionate ratio and Firmicutes to Bacteroidetes ratio.
```

### Homework 6.5
Import the curated pH+BSS data you generated during Lab4. Join the pH data frame with the tax_ratio data frame.
```{r}
# import 
pH_BSS_wkly <- read_delim(file = "Lab4/curated_data/pH_BSS_wkly.csv",
                          delim = "\t", col_names = TRUE, trim_ws = TRUE)

# join 
tax_scfa_pH <- inner_join(tax_ratio, pH_BSS_wkly) %>%
  na.omit(semester)
```

Determine if there is a relationship between total SCFAs and acetate. Conduct the statistical test and generate a plot. Save the plot to figures folder.
```{r}
scfa_acetate_plot <- tax_ratio %>%
  ggplot(aes(x = totalscfa_mean,
             y = acetate_mean)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE) 

scfa_acetate_plot

save_plot(filename = "Lab6/figures/scfa_acetate_plot.pdf",
          plot = scfa_acetate_plot,
          nrow = 1, ncol = 2,
          base_aspect_ratio = 1.1)

m63e <- tax_ratio %>%
  lm(totalscfa_mean ~ acetate_mean, data = .) 

summary(m63e) # There is very clearly a strong linear relationship between total SCFAs and acetate. The p-value is low, and looking at the adjusted R-squared value (0.9114) it is evident that the linear relationship is a good fit.
```

Do you draw the same conclusions if you use the total SCFAs and acetate ratio?
```{r}
scfa_ace_ratio_plot <- tax_ratio %>%
  ggplot(aes(x = totalscfa_mean,
             y = acetate_ratio)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE) 

scfa_ace_ratio_plot

m63f <- tax_ratio %>%
  lm(totalscfa_mean ~ acetate_ratio, data = .) 

summary(m63f) # The relationship is not as clear when comparing total SCFAs and acetate ratio. Although there seems to be a very slight upward trend in the data points shown by the line of best fit, we have an adjusted R-squared value of 0.05, indicating a very weak linear relationship.

```

### Homework 6.6
Determine if there is a relationship between total SCFAs and pH. Conduct the statistical test and generate a plot. Save the plot.
```{r}
scfa_pH_plot <- tax_scfa_pH %>%
  ggplot(aes(x = totalscfa_mean,
             y = avg_wkly_pH)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE)

scfa_pH_plot

save_plot(filename = "Lab6/figures/scfa_pH_plot.pdf",
          plot = scfa_pH_plot,
          nrow = 1, ncol = 2,
          base_aspect_ratio = 1.1)

m63g <- tax_scfa_pH %>%
  lm(totalscfa_mean ~ avg_wkly_pH, data = .) 

summary(m63g) # Although a low p-value was produced, looking at the adjusted R-squared value it is evident that the linear relationship between total SCFAs and pH is weak, because the value is relatively low (0.1). There is a very slight negative linear relationship, if any.

```

### Homework 6.7
Determine if there is a relationship between butyrate and pH. Conduct the statistical test and generate a plot. Save the plot.
```{r}
butyrate_pH_plot <- tax_scfa_pH %>%
  ggplot(aes(x = butyrate_mean,
             y = avg_wkly_pH)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = FALSE) 

butyrate_pH_plot

save_plot(filename = "Lab6/figures/butyrate_pH_plot.pdf",
          plot = butyrate_pH_plot,
          nrow = 1, ncol = 2,
          base_aspect_ratio = 1.1)

m63h <- tax_scfa_pH %>%
  lm(butyrate_mean ~ avg_wkly_pH, data = .) 

summary(m63h) # The p-value generated by this test is low, but the adjusted R-squared value (0.05) indicates no strong linear relationship between butyrate and pH. The line of best fit indicates a slightly negative linear trend, but regardless the data is scattered and does not very clearly show a pattern.

```

### Homework 6.8
Save each data frame generated today:

* taxa_cat_curated
* scfa_ratio
* tax_scfa
* tax_scfa_pH

```{r}
write_delim(taxa_cat_curated, 
            path="Lab6/curated_data/taxa_cat_curated.txt",
            delim = "\t", quote = FALSE, col_names = TRUE)

write_delim(scfa_ratio, 
            path="Lab6/curated_data/scfa_ratio.txt",
            delim = "\t", quote = FALSE, col_names = TRUE)

write_delim(tax_scfa, 
            path="Lab6/curated_data/tax_scfa.txt",
            delim = "\t", quote = FALSE, col_names = TRUE)

write_delim(tax_scfa_pH, 
            path="Lab6/curated_data/tax_scfa_pH.txt",
            delim = "\t", quote = FALSE, col_names = TRUE)
```

-----
end